---
title: "Introduction"
output: 
  bookdown::html_vignette2:
bibliography: '`r system.file("tmap.bib", package="tmap")`'
csl: "`r system.file('ieee.csl', package = 'tmap')`"
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  results = 'hide',
  fig.width = 8,
  comment = "#>"
)
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
   lines <- options$output.lines
   if (is.null(lines)) {
     return(hook_output(x, options))  # pass to default hook
   }
   x <- unlist(strsplit(x, "\n"))
   more <- "..."
   if (length(lines)==1) {        # first n lines
     if (length(x) > lines) {
       # truncate the output, but add ....
       x <- c(head(x, lines), more)
     }
   } else {
     x <- c(more, x[lines], more)
   }
   # paste these lines together
   x <- paste(c(x, ""), collapse = "\n")
   hook_output(x, options)
 })

```


```{r, echo = FALSE, message = FALSE}
library(tmap)
library(tmap.cartogram)
tmap_options(scale = 0.75)
```



## Cartogram

In cartograms, areal regions are distorted such that the obtained area sizes are proportional to a quantitative variable. We facilitate the following types:

* The **contiguous cartogram**, where the shapes of the polygons are distorted, where the neighborhood relationships between them are preserved as much as possible.
* The **non-contiguous cartogram**, where the polygons are only resized while they keep their shape.
* The **Dorling cartogram** where polygons are replaced by bubbles. Note that this is different from the [bubble map](https://r-tmap.github.io/tmap/articles/examples_bubble): in a Dorlong cartogram the bubbles do not overlap. To achieve non-overlapping bubbles that are proportional to the target variable, the bubbles may be relocated.

Currently, this packages uses the [cartogram](https://sjewo.github.io/cartogram/) package by Sebastian Jeworutzki under the hood.

## Map projection

In order to work well, it is strongly recommended to use a projected [CRS (map projection)](https://r-tmap.github.io/tmap/articles/foundations_crs) that is (approximately) equal-area. Why? Because the cartograms transformation require that that the area of polygons is approximately proportional to the real areas. This is not the case for the mercator projection (plain lat/lon coordinates).

For these examples we use the Robinson projection. If you are unsure which to use, set `crs = "auto"` which enables a (work-in-progress) automatic map projection recommendation.

In addition, this crs must be set in [`tm_shape()`](https://r-tmap.github.io/tmap/reference/tm_shape.html), and not afterwards via [`tm_crs()`](https://r-tmap.github.io/tmap/reference/tm_crs.html). Why? Because the former ensures that the transformation is applied in the correct CRS, whereas the latter is used for plotting only.

Finally, we recommend to disable basemaps in view mode. Not only because the polygons will be distorted (and therefore will not be consistent with the basemaps), but because basemaps are usually rendered in the (pseudo) mercator projection. By default a spatial object will keep its projected CRS when basemaps are disabled.
